---
title: "Morley"
author: Ben Tupper
date: April 8, 2020
output: md_document
---

## Replicate the examples found in `?morley` using tidy tools

Here are the steps and output of the examples found in `?morley`.

```{r}
require(stats)
require(graphics)
michelson <- transform(morley,
                       Expt = factor(Expt), Run = factor(Run))
xtabs(~ Expt + Run, data = michelson)  # 5 x 20 balanced (two-way)
```

```{r}
plot(Speed ~ Expt, data = michelson,
     main = "Speed of Light Data", xlab = "Experiment No.")
```

```{r}
fm <- aov(Speed ~ Run + Expt, data = michelson)
summary(fm)
```

```{r}
fm0 <- update(fm, . ~ . - Run)
anova(fm0, fm)
```

### Challenges

Start by casting the `morley` data as a tibble rather than as a data.frame.

```{r}
library(dplyr)
x <- morley %>%
  dplyr::as_tibble()
```


Q1: How might you produce the contignecy table (2d counts) using tidy tools?

Q2: How might you generate the figure using ggplot2 and friends? (Bonus, can you make it a violin plot?)

Q3: How might you use tidy tools to extract coefficients and metrics out of the models `fm` and `fm0`?


